---
title: An R Markdown document converted from "scaling_reference_seperately_and_then_performing_integration.ipynb"
output: html_document
---

  - As an alternative to try and improve the integration of the two datasets, first perform QC, scaling and normalization seperately.
  - Then merge both objects into one.
  - Perform scVI integration using batch_key=
      - Sample
      - Technology

## Load the Required Packages:

```{r}
setwd("/home/maryam/Desktop/scVI/integrating_marieke_data_with_tsp/")

options(repr.plot.width=12, repr.plot.height=6)

options(scipen=100) #avoid scientific notation of numbers

library(SeuratDisk)
library(Seurat)
library(matrixStats)
library(ggplot2)
library(pheatmap)
library(reshape2)
library(dplyr)
library(tidyr)
library(viridis)
library(SingleCellExperiment)
library(tidyverse)
library(Matrix)
library(scales)
library(cowplot)
library(RCurl)
library(SeuratDisk)
library(Seurat)
library(matrixStats)
library(ggplot2)
library(pheatmap)
library(reshape2)
library(dplyr)
library(tidyr)
library(viridis)
library(batchelor)
library(BiocParallel)
library(BiocNeighbors)

# Packages needed for scVI:
library(sceasy)
library(reticulate)
```

## Creating the Seperate Seurat Objects:

### Seurat Object for Marieke's dataset:

```{r}
h5_CISE10 <- list.files(path="/home/maryam/Desktop/marieke_data/CISE10/",  #path to file folder, relative to wd
             recursive=TRUE,  #also include potential subdirectories
             pattern=".h5",  #return files matching this expression
             full.names=TRUE) #get full path instead of just filename
h5_CISE10

h5_CISE11 <- list.files(path="/home/maryam/Desktop/marieke_data/CISE11/",  #path to file folder, relative to wd
             recursive=TRUE,  #also include potential subdirectories
             pattern=".h5",  #return files matching this expression
             full.names=TRUE) #get full path instead of just filename
h5_CISE11

h5_S1 <- list.files(path="/home/maryam/Desktop/marieke_data/S1/",  #path to file folder, relative to wd
         recursive=TRUE,  #also include potential subdirectories
         pattern=".h5",  #return files matching this expression
         full.names=TRUE) #get full path instead of just filename
h5_S1

h5_S2 <- list.files(path="/home/maryam/Desktop/marieke_data/S2/",  #path to file folder, relative to wd
         recursive=TRUE,  #also include potential subdirectories
         pattern=".h5",  #return files matching this expression
         full.names=TRUE) #get full path instead of just filename
h5_S2

h5_S3 <- list.files(path="/home/maryam/Desktop/marieke_data/S3/",  #path to file folder, relative to wd
         recursive=TRUE,  #also include potential subdirectories
         pattern=".h5",  #return files matching this expression
         full.names=TRUE) #get full path instead of just filename
h5_S3

h5_TTA15 <-  list.files(path="/home/maryam/Desktop/marieke_data/TTA15/",  #path to file folder, relative to wd
             recursive=TRUE,  #also include potential subdirectories
             pattern=".h5",  #return files matching this expression
             full.names=TRUE) #get full path instead of just filename
h5_TTA15

h5_TTA16 <-  list.files(path="/home/maryam/Desktop/marieke_data/TTA16/",  #path to file folder, relative to wd
             recursive=TRUE,  #also include potential subdirectories
             pattern=".h5",  #return files matching this expression
             full.names=TRUE) #get full path instead of just filename
h5_TTA16
```

```{r}
# Create a Seurat object for each sample
#CISE10:
for (file in c("CISE10/")){
        cise10_seurat_data <- Read10X(data.dir = paste0("", file))
        cise10_seurat_obj <- CreateSeuratObject(counts = cise10_seurat_data, 
                                         min.features = 10, 
                                         project = file)
        assign(file, cise10_seurat_obj)
}


#CISE11:
for (file in c("CISE11/")){
        cise11_seurat_data <- Read10X(data.dir = paste0("", file))
        cise11_seurat_obj <- CreateSeuratObject(counts = cise11_seurat_data, 
                                         min.features = 10, 
                                         project = file)
        assign(file, cise11_seurat_obj)
}

#S1:
for (file in c("S1/")){
        s1_seurat_data <- Read10X(data.dir = paste0("", file))
        s1_seurat_obj <- CreateSeuratObject(counts = s1_seurat_data, 
                                         min.features = 10, 
                                         project = file)
        assign(file, s1_seurat_obj)
}

#S2:
for (file in c("S2/")){
        s2_seurat_data <- Read10X(data.dir = paste0("", file))
        s2_seurat_obj <- CreateSeuratObject(counts = s2_seurat_data, 
                                         min.features = 10, 
                                         project = file)
        assign(file, s2_seurat_obj)
}


#S3:
for (file in c("S3/")){
        s3_seurat_data <- Read10X(data.dir = paste0("", file))
        s3_seurat_obj <- CreateSeuratObject(counts = s3_seurat_data, 
                                         min.features = 10, 
                                         project = file)
        assign(file, s3_seurat_obj)
}

#TTA15:
for (file in c("TTA15//")){
        tta15_seurat_data <- Read10X(data.dir = paste0("", file))
        tta15_seurat_obj <- CreateSeuratObject(counts = tta15_seurat_data, 
                                         min.features = 10, 
                                         project = file)
        assign(file, tta15_seurat_obj)
}

#TTA16:
for (file in c("TTA16//")){
        tta16_seurat_data <- Read10X(data.dir = paste0("", file))
        tta16_seurat_obj <- CreateSeuratObject(counts = tta16_seurat_data, 
                                         min.features = 10, 
                                         project = file)
        assign(file, tta16_seurat_obj)
}
```

#### Merging the above Samples into one Seurat Object:

```{r}
#Merging all the Samples into one Seurat Object:
marieke_seurat <- merge(x = cise10_seurat_obj, 
                      y = c(cise11_seurat_obj, s1_seurat_obj, s2_seurat_obj, s3_seurat_obj, tta15_seurat_obj, tta16_seurat_obj),
                      add.cell.id = c("cise10", "cise11", "s1", "s2", "s3", "tta15", "tta16"))
```

```{r}
marieke_seurat
```

```{r}
head(marieke_seurat)
```

### Seurat Object for the TSP dataset:

```{r}
h5_MO_TSP1 <- list.files(path="/home/maryam/Desktop/Multiome_GEX_data/MO-TSP1_GEX1/",  #path to file folder, relative to wd
             recursive=TRUE,  #also include potential subdirectories
             pattern=".h5",  #return files matching this expression
             full.names=TRUE) #get full path instead of just filename

h5_MO_TSP1

h5_MO_TSP3 <- list.files(path="/home/maryam/Desktop/Multiome_GEX_data/MO-TSP3_GEX3/",  #path to file folder, relative to wd
             recursive=TRUE,  #also include potential subdirectories
             pattern=".h5",  #return files matching this expression
             full.names=TRUE) #get full path instead of just filename

h5_MO_TSP3

h5_MO_TSP4 <- list.files(path="/home/maryam/Desktop/Multiome_GEX_data/MO-TSP4_GEX4/",  #path to file folder, relative to wd
             recursive=TRUE,  #also include potential subdirectories
             pattern=".h5",  #return files matching this expression
             full.names=TRUE) #get full path instead of just filename

h5_MO_TSP4
```

```{r}
#Donor1:
for (file in c("MO-TSP1_GEX1/")){
        motsp1_seurat_data <- Read10X(data.dir = paste0("", file))
        motsp1_seurat_obj <- CreateSeuratObject(counts = motsp1_seurat_data, 
                                         min.features = 10, 
                                         project = file)
        assign(file, motsp1_seurat_obj)
}


#Donor3:
for (file in c("MO-TSP3_GEX3/")){
        motsp3_seurat_data <- Read10X(data.dir = paste0("", file))
        motsp3_seurat_obj <- CreateSeuratObject(counts = motsp3_seurat_data, 
                                         min.features = 10, 
                                         project = file)
        assign(file, motsp3_seurat_obj)
}

#Donor4:
for (file in c("MO-TSP4_GEX4/")){
        motsp4_seurat_data <- Read10X(data.dir = paste0("", file))
        motsp4_seurat_obj <- CreateSeuratObject(counts = motsp4_seurat_data, 
                                         min.features = 10, 
                                         project = file)
        assign(file, motsp4_seurat_obj)
}
```

#### Merging the above Samples into one Seurat Object:

```{r}
#Merging all the Samples into one Seurat Object:
tsp_seurat <- merge(x = motsp1_seurat_obj, 
                      y = c(motsp3_seurat_obj, motsp4_seurat_obj),
                      add.cell.id = c("motsp1", "motsp3", "motsp4"))
```

```{r}
tsp_seurat
```

```{r}
head(tsp_seurat@meta.data)
```

### Creating Sample and Technology Columns:

#### Marieke's dataset:

```{r}
# Create metadata dataframe
marieke_metadata <- marieke_seurat@meta.data

# Add cell IDs to metadata
marieke_metadata$cells <- rownames(marieke_metadata)

# Create sample column
marieke_metadata$sample <- NA
marieke_metadata$sample[which(str_detect(marieke_metadata$cells, "^cise10_"))] <- "marieke_cise10"
marieke_metadata$sample[which(str_detect(marieke_metadata$cells, "^cise11_"))] <- "marieke_cise11"
marieke_metadata$sample[which(str_detect(marieke_metadata$cells, "^s1_"))] <- "marieke_s1"
marieke_metadata$sample[which(str_detect(marieke_metadata$cells, "^s2_"))] <- "marieke_s2"
marieke_metadata$sample[which(str_detect(marieke_metadata$cells, "^s3_"))] <- "marieke_s3"
marieke_metadata$sample[which(str_detect(marieke_metadata$cells, "^tta15_"))] <- "marieke_tta15"
marieke_metadata$sample[which(str_detect(marieke_metadata$cells, "^tta16_"))] <- "marieke_tta16"

# Create technology column:
marieke_metadata$technology <- NA
marieke_metadata$technology[which(str_detect(marieke_metadata$cells, "^cise10_"))] <- "total_RNA"
marieke_metadata$technology[which(str_detect(marieke_metadata$cells, "^cise11_"))] <- "total_RNA"
marieke_metadata$technology[which(str_detect(marieke_metadata$cells, "^s1_"))] <- "total_RNA"
marieke_metadata$technology[which(str_detect(marieke_metadata$cells, "^s2_"))] <- "total_RNA"
marieke_metadata$technology[which(str_detect(marieke_metadata$cells, "^s3_"))] <- "total_RNA"
marieke_metadata$technology[which(str_detect(marieke_metadata$cells, "^tta15_"))] <- "total_RNA"
marieke_metadata$technology[which(str_detect(marieke_metadata$cells, "^tta16_"))] <- "total_RNA"

# Create donor_ID column:
marieke_metadata$donor_id <- NA
marieke_metadata$donor_id[which(str_detect(marieke_metadata$cells, "^cise10_"))] <- "ML_A"
marieke_metadata$donor_id[which(str_detect(marieke_metadata$cells, "^cise11_"))] <- "ML_B"
marieke_metadata$donor_id[which(str_detect(marieke_metadata$cells, "^s1_"))] <- "ML_A"
marieke_metadata$donor_id[which(str_detect(marieke_metadata$cells, "^s2_"))] <- "ML_B"
marieke_metadata$donor_id[which(str_detect(marieke_metadata$cells, "^s3_"))] <- "ML_C"
marieke_metadata$donor_id[which(str_detect(marieke_metadata$cells, "^tta15_"))] <- "ML_E"
marieke_metadata$donor_id[which(str_detect(marieke_metadata$cells, "^tta16_"))] <- "ML_F"

# Create kit column:
marieke_metadata$kit <- NA
marieke_metadata$kit[which(str_detect(marieke_metadata$cells, "^cise10_"))] <- "sc3'v2"
marieke_metadata$kit[which(str_detect(marieke_metadata$cells, "^cise11_"))] <- "sc3'v2"
marieke_metadata$kit[which(str_detect(marieke_metadata$cells, "^s1_"))] <- "sc3'v2"
marieke_metadata$kit[which(str_detect(marieke_metadata$cells, "^s2_"))] <- "sc3'v2"
marieke_metadata$kit[which(str_detect(marieke_metadata$cells, "^s3_"))] <- "sc3'v2"
marieke_metadata$kit[which(str_detect(marieke_metadata$cells, "^tta15_"))] <- "sc3'v2"
marieke_metadata$kit[which(str_detect(marieke_metadata$cells, "^tta16_"))] <- "sc3'v2"
```

```{r}
marieke_metadata
```

```{r}
## Add Metadata back to Seurat Object:
marieke_seurat@meta.data <- marieke_metadata
View(marieke_seurat@meta.data)
```

#### MO-TSP dataset:

```{r}
# Create metadata dataframe
motsp_metadata <- tsp_seurat@meta.data

# Add cell IDs to metadata
motsp_metadata$cells <- rownames(motsp_metadata)

# Create sample column
motsp_metadata$sample <- NA
motsp_metadata$sample[which(str_detect(motsp_metadata$cells, "^motsp1_"))] <- "motsp1"
motsp_metadata$sample[which(str_detect(motsp_metadata$cells, "^motsp3_"))] <- "motsp3"
motsp_metadata$sample[which(str_detect(motsp_metadata$cells, "^motsp4_"))] <- "motsp4"

# Create technology column:
motsp_metadata$technology <- NA
motsp_metadata$technology[which(str_detect(motsp_metadata$cells, "^motsp1_"))] <- "nuclei_RNA"
motsp_metadata$technology[which(str_detect(motsp_metadata$cells, "^motsp3_"))] <- "nuclei_RNA"
motsp_metadata$technology[which(str_detect(motsp_metadata$cells, "^motsp4_"))] <- "nuclei_RNA"

# Create donor_id column:
motsp_metadata$donor_id <- NA
motsp_metadata$donor_id[which(str_detect(motsp_metadata$cells, "^motsp1_"))] <- "ND_04"
motsp_metadata$donor_id[which(str_detect(motsp_metadata$cells, "^motsp3_"))] <- "TSC_CT14"
motsp_metadata$donor_id[which(str_detect(motsp_metadata$cells, "^motsp4_"))] <- "TSC_CT17"

# Create kit column:
motsp_metadata$kit <- NA
motsp_metadata$kit[which(str_detect(motsp_metadata$cells, "^motsp1_"))] <- "multiome"
motsp_metadata$kit[which(str_detect(motsp_metadata$cells, "^motsp3_"))] <- "multiome"
motsp_metadata$kit[which(str_detect(motsp_metadata$cells, "^motsp4_"))] <- "multiome"
```

```{r}
head(motsp_metadata)
```

```{r}
## Add Metadata back to the Seurat Object: 
tsp_seurat@meta.data <- motsp_metadata
View(tsp_seurat@meta.data)
```

## Quality Control:
### Marieke dataset:

```{r}
## Calculating the percent.mt and visualizing the distribution:
marieke_seurat [["percent.mt"]] <- PercentageFeatureSet(marieke_seurat, pattern = "MT")
marieke_data <- marieke_seurat@meta.data
head(marieke_data)
```

```{r}
plot_1 <- ggplot(marieke_data, aes(x=log10(nCount_RNA), y=percent.mt, colour=sample)) +
          geom_point(size=0.05) +
          facet_wrap(~interaction(orig.ident,sample), ncol=5)+
          labs(title = "Counts vs. mito%") +
          scale_colour_viridis_d()+
          geom_hline(yintercept = 20) +
          theme_bw()

options(repr.plot.width=14, repr.plot.height=10)
plot_1
```

```{r}
marieke_seurat$keep_viab <- !marieke_seurat$percent.mt >=20  
marieke_data2 <- marieke_seurat@meta.data                     
marieke_data2 <- marieke_data2[order(marieke_data2$keep, decreasing=FALSE),]
```

```{r}
plot_2 <- ggplot(marieke_data2, aes(x=log10(nCount_RNA), y=percent.mt, colour=keep_viab))+
            geom_point(size=0.03, alpha=0.5)+
            facet_wrap(~interaction(orig.ident, sample), ncol=5)+
            labs(title='Counts vs. mito%')+
            theme_bw()

plot_3 <- ggplot(marieke_data2, aes(x=log10(nCount_RNA), y=log10(nFeature_RNA), colour=keep_viab))+
            geom_point(size=0.03, alpha=0.5)+
            facet_wrap(~interaction(orig.ident, sample), ncol=5)+
            geom_hline(yintercept = 20) +
            labs(title='Counts vs. Features')+
            theme_bw()

options(repr.plot.width=14, repr.plot.height=10)
plot_2 
```

```{r}
## Identifying cells with low read/gene count which may correspond to empty droplets:

options(repr.plot.width=14, repr.plot.height=10)

ggplot(marieke_data, aes(x=log10(nCount_RNA), y=log10(nFeature_RNA), colour=sample))+
            geom_point(size=0.05)+
            facet_wrap(~interaction(orig.ident, sample), ncol = 5)+
            labs(title='Count vs Features')+
            scale_color_viridis_d()+
            theme_bw()
```

```{r}
marieke_seurat$keep_counts <- !c(marieke_seurat$nFeature_RNA <= 500 | marieke_seurat$nCount_RNA <=1000)
```

```{r}
options(repr.plot.width=18, repr.plot.height=10)

plot_4 <- ggplot(marieke_seurat@meta.data, aes(x=log10(nCount_RNA), y=log10(nFeature_RNA), colour=keep_counts))+
            geom_point(size=0.03, alpha=0.5)+ #(change 0.03 to 1 for larger dot)
            facet_wrap(~interaction(orig.ident, sample), ncol=5)+
            labs(title='Counts vs. Features')+
            theme_bw()


plot_4
```

```{r}
## Remove the low quality cells:

marieke_seurat@meta.data$keep <- marieke_seurat@meta.data$keep_counts & marieke_seurat@meta.data$keep_viab

table(marieke_seurat@meta.data$keep)
marieke_seurat@meta.data %>%
group_by(orig.ident,sample) %>%
summarise("total cells"=length(keep), "remaining"=sum(keep), "remove"=(length(keep)-sum(keep)), "% remove"=round(((length(keep)-sum(keep))/length(keep)*100),1))
```

```{r}
marieke_seurat <- subset(marieke_seurat, subset=keep_counts & keep_viab)
dim(marieke_seurat)
```

```{r}
## Removing the non-informative genes:

keep_genes <- rowSums(marieke_seurat@assays$RNA@counts>0 ) > 3
table(keep_genes)
```

```{r}
marieke_seurat <- marieke_seurat[keep_genes,]
dim(marieke_seurat)
```

#### Doublet Removal for Marieke's Dataset:

```{r}
library(scDblFinder)
library(BiocParallel)
```

```{r}
marieke_sce <- as.SingleCellExperiment(marieke_seurat)
```

```{r}
start.time <- Sys.time()

marieke_sce <- scDblFinder(marieke_sce, samples = marieke_seurat@meta.data$filename)
table(sce$scDblFinder.class)

end.time <- Sys.time()
time.taken <- end.time - start.time
time.taken
```

```{r}
table(marieke_sce$scDblFinder.class)

end.time <- Sys.time()
time.taken <- end.time - start.time
time.taken
```

```{r}
marieke_doublets <- rownames(colData(marieke_sce)[marieke_sce$scDblFinder.class=='doublet',])
marieke_seurat$singlet <- NA

marieke_seurat$singlet <-ifelse(rownames(marieke_seurat@meta.data) %in% marieke_doublets, 'FALSE', 'TRUE')
```

```{r}
options(repr.plot.width=14, repr.plot.height=10)

plot_5 <- ggplot(marieke_seurat@meta.data, aes(x=log10(nCount_RNA), y=log10(nFeature_RNA), colour=singlet))+
            geom_point(size=0.03, alpha=0.5)+
            facet_wrap(~interaction(orig.ident, sample), ncol=5)+
            labs(title='Counts vs. Features')+
            theme_bw()


plot_5
```

```{r}
marieke_seurat$singlet <- as.logical(marieke_seurat$singlet)
marieke_seurat <- subset(marieke_seurat, subset=singlet)
dim(marieke_seurat)
```

#### Performing Normalization and Scaling on Marieke's Dataset:

```{r}
marieke_seurat
```

```{r}
marieke_seurat <- NormalizeData(marieke_seurat, normalization.method = "LogNormalize", scale.factor = 10000, assay = "RNA")
```

```{r}
marieke_seurat <- ScaleData(marieke_seurat)
```

### MO-TSP Dataset:

```{r}
## Calculating the percent.mt and visualizing the distribution:
tsp_seurat [["percent.mt"]] <- PercentageFeatureSet(tsp_seurat, pattern = "MT")
motsp_data <- tsp_seurat@meta.data
head(motsp_data)
```

```{r}
plot_a <- ggplot(motsp_data, aes(x=log10(nCount_RNA), y=percent.mt, colour=sample)) +
          geom_point(size=0.05) +
          facet_wrap(~interaction(orig.ident,sample), ncol=5)+
          labs(title = "Counts vs. mito%") +
          scale_colour_viridis_d()+
          geom_hline(yintercept = 20) +
          theme_bw()

options(repr.plot.width=14, repr.plot.height=10)
plot_a
```

```{r}
tsp_seurat$keep_viab <- !tsp_seurat$percent.mt >=20  
motsp_data2 <- tsp_seurat@meta.data                     
motsp_data2 <- motsp_data2[order(motsp_data2$keep, decreasing=FALSE),]
```

```{r}
plot_b <- ggplot(motsp_data2, aes(x=log10(nCount_RNA), y=percent.mt, colour=keep_viab))+
            geom_point(size=0.03, alpha=0.5)+
            facet_wrap(~interaction(orig.ident, sample), ncol=5)+
            labs(title='Counts vs. mito%')+
            theme_bw()

plot_c <- ggplot(motsp_data2, aes(x=log10(nCount_RNA), y=log10(nFeature_RNA), colour=keep_viab))+
            geom_point(size=0.03, alpha=0.5)+
            facet_wrap(~interaction(orig.ident, sample), ncol=5)+
            geom_hline(yintercept = 20) +
            labs(title='Counts vs. Features')+
            theme_bw()

options(repr.plot.width=14, repr.plot.height=10)
plot_b 
```

```{r}
## Identifying cells with low read/gene count which may correspond to empty droplets:

options(repr.plot.width=14, repr.plot.height=10)

ggplot(motsp_data, aes(x=log10(nCount_RNA), y=log10(nFeature_RNA), colour=sample))+
            geom_point(size=0.05)+
            facet_wrap(~interaction(orig.ident, sample), ncol = 5)+
            labs(title='Count vs Features')+
            scale_color_viridis_d()+
            theme_bw()
```

```{r}
tsp_seurat$keep_counts <- !c(tsp_seurat$nFeature_RNA <= 270 | tsp_seurat$nCount_RNA <=700)
```

```{r}
options(repr.plot.width=18, repr.plot.height=10)

plot_d <- ggplot(tsp_seurat@meta.data, aes(x=log10(nCount_RNA), y=log10(nFeature_RNA), colour=keep_counts))+
            geom_point(size=0.03, alpha=0.5)+ #(change 0.03 to 1 for larger dot)
            facet_wrap(~interaction(orig.ident, sample), ncol=5)+
            labs(title='Counts vs. Features')+
            theme_bw()


plot_d
```

```{r}
## Remove the low quality cells:

tsp_seurat@meta.data$keep <- tsp_seurat@meta.data$keep_counts & tsp_seurat@meta.data$keep_viab

table(tsp_seurat@meta.data$keep)
tsp_seurat@meta.data %>%
group_by(orig.ident,sample) %>%
summarise("total cells"=length(keep), "remaining"=sum(keep), "remove"=(length(keep)-sum(keep)), "% remove"=round(((length(keep)-sum(keep))/length(keep)*100),1))
```

```{r}
tsp_seurat <- subset(tsp_seurat, subset=keep_counts & keep_viab)
dim(tsp_seurat)
```

```{r}
## Removing the non-informative genes:

keep_genes <- rowSums(tsp_seurat@assays$RNA@counts>0 ) > 3
table(keep_genes)
```

```{r}
tsp_seurat <- tsp_seurat[keep_genes,]
dim(tsp_seurat)
```

```{r}
#### Doublet Removal for MO-TSP Dataset:
```

```{r}
motsp_sce <- as.SingleCellExperiment(tsp_seurat)
```

```{r}
start.time <- Sys.time()

motsp_sce <- scDblFinder(motsp_sce, samples = tsp_seurat@meta.data$filename)
table(motsp_sce$scDblFinder.class)

end.time <- Sys.time()
time.taken <- end.time - start.time
time.taken
```

```{r}
motsp_doublets <- rownames(colData(motsp_sce)[motsp_sce$scDblFinder.class=='doublet',])
tsp_seurat$singlet <- NA

tsp_seurat$singlet <-ifelse(rownames(tsp_seurat@meta.data) %in% motsp_doublets, 'FALSE', 'TRUE')
```

```{r}
options(repr.plot.width=14, repr.plot.height=10)

plot_e <- ggplot(tsp_seurat@meta.data, aes(x=log10(nCount_RNA), y=log10(nFeature_RNA), colour=singlet))+
            geom_point(size=0.03, alpha=0.5)+
            facet_wrap(~interaction(orig.ident, sample), ncol=5)+
            labs(title='Counts vs. Features')+
            theme_bw()


plot_e
```

```{r}
tsp_seurat$singlet <- as.logical(tsp_seurat$singlet)
tsp_seurat <- subset(tsp_seurat, subset=singlet)
dim(tsp_seurat)
```

#### Performing Normalization and Scaling on MO-TSP Dataset:

```{r}
tsp_seurat
```

```{r}
tsp_seurat <- NormalizeData(tsp_seurat, normalization.method = "LogNormalize", scale.factor = 10000, assay = "RNA")
```

```{r}
tsp_seurat <- ScaleData(tsp_seurat)
```

```{r}
marieke_seurat
```

```{r}
tsp_seurat
```

## Merge the two Seurat Objects into one:

```{r}
merged_seurat <- merge(x=marieke_seurat, y= tsp_seurat)
```

```{r}
merged_seurat
```

```{r}
head(merged_seurat@meta.data)
```

```{r}
tail(merged_seurat@meta.data)
```

```{r}
seurObj <- merged_seurat
```

## Find HVG's and Perform Integration:

```{r}
## Finding Highly Variable Genes:
seurObj<- FindVariableFeatures(seurObj, selection.method = "vst", nfeatures = 2000)
top2000 <- head(VariableFeatures(seurObj), 2000)
seurObj <- seurObj[top2000]
seurObj
```

```{r}
## Save this new seurat object:
SaveH5Seurat(seurObj, "seperately_")
```

