---
title: An R Markdown document converted from "Normalisation and Scaling - Part 2.ipynb"
output: html_document
---

## Loading the Necessary Packages:

```{r}
setwd("/home/maryam/Desktop/Multiome_GEX_data")

options(repr.plot.width=12, repr.plot.height=6)

options(scipen=100) #avoid scientific notation of numbers

library(SeuratDisk)
library(Seurat)
library(matrixStats)
library(ggplot2)
library(pheatmap)
library(reshape2)
library(dplyr)
library(tidyr)
library(viridis)
library(SingleCellExperiment)
library(tidyverse)
library(Matrix)
library(scales)
library(cowplot)
library(RCurl)
```

## Loading Our Data:

```{r}
seurObj <- LoadH5Seurat("GEX_QC.h5seurat")
```

```{r}
View(seurObj)
```

### Determining Cell Cycle Events:

  - The Seurat package comes with a list of cell cycle genes, which can be used to identify and regress out related effects.

```{r}
s.genes <- cc.genes.updated.2019$s.genes
g2m.genes <- cc.genes.updated.2019$g2m.genes
```

  - Before determining the cell cycle score, data is supposed to be normalised acc. to https://github.com/satijalab/seurat/issues/1679

```{r}
seurObj <- NormalizeData(seurObj, normalization.method = "LogNormalize", scale.factor=10000, assay = 'RNA')

seurObj <- FindVariableFeatures(seurObj, selection.method = "vst", nfeatures = 2000, assay = 'RNA')
```

  - Cell cycle scoring adds a S and G2M score and the predicted cell cycle phase to the cell meta data.

```{r}
seurObj <- CellCycleScoring(seurObj, s.features = s.genes, g2m.features = g2m.genes, set.ident = TRUE)
```

```{r}
seurObj$CC.Difference <- seurObj$S.Score - seurObj$G2M.Score
```

```{r}
ptm <- proc.time()

seurObj <- ScaleData(seurObj, assay = 'RNA', use.umi = TRUE)

proc.time() - ptm
```

```{r}
ptm <- proc.time()

seurObj <- RunPCA(seurObj, features = VariableFeatures(object = seurObj), reduction.name='pca') #RunPCA automatically operates on the scale.data slot

proc.time() - ptm
```

```{r}
options(repr.plot.width=20, repr.plot.height=10)

p1 <- DimPlot(seurObj, reduction='pca', shuffle=T, group.by = 'Phase', split.by='sample') + labs(title='No cell cycle correction')

p1
```

  - From the above, we can see that the PCA confirms that the cells cluster by cell cycle phase.

  - We can also compare the top genes in the first PCs with the known cell cycle genes.

```{r}
options(repr.plot.width=15, repr.plot.height=15)

DimHeatmap(seurObj, dims=c(1:4), reduction= 'pca', cells=1000, balanced=TRUE, ncol=2, fast=FALSE) &
scale_fill_viridis(direction = 1)
```

```{r}
#Extract genes with highest impact on PC

PCA_features <- c() #create empty vector the build data frame from 

for (i in 1:5){   #iterate over first five PCs
    loadings <- data.frame(seurObj@reductions$pca@feature.loadings[,i, drop=F]) #get all loadings for the PC
    up <- loadings %>% top_n(n=15) %>% rownames()  #get the 15 genes with highest loadings
    down <- loadings %>% top_n(n=-15) %>% rownames() #get the 15 genes with loasted loadings
    PCA_features <- cbind(PCA_features, c(up, down)) #add to data frame
    colnames(PCA_features)[i] <- paste0('PC', i)  #name column by PC
    }
```

  - Could the top most influential genes per PC be known cell cycle genes?

```{r}
PCA_cc.genes <- c()
for (i in 1:ncol(PCA_features)){
    PCA_cc.genes <- c(PCA_cc.genes, sum(PCA_features[,i] %in% s.genes | PCA_features[,i] %in% g2m.genes))
    names(PCA_cc.genes)[i] <- paste0('PC', i)
    }

PCA_cc.genes
```

```{r}
sort(s.genes)
sort(g2m.genes)
```

  - We can regress repeat the scaling and regress out the cell cycle difference. In this instance we also set use.umi to TRUE, which will result in libray size correction. This is because we will use the corrected data for further analysis and would like to account for this potential technical variation.

```{r}
ptm <- proc.time()

seurObj <- ScaleData(seurObj, assay = 'RNA', vars.to.regress = c('CC.Difference'), use.umi = TRUE)

proc.time() - ptm
```

```{r}
ptm <- proc.time()

seurObj <- RunPCA(seurObj, features = VariableFeatures(object = seurObj), reduction.name='pca_CC') #RunPCA automatically operates on the scale.data slot

proc.time() - ptm
```

```{r}
options(repr.plot.width=20, repr.plot.height=6)

p2 <- DimPlot(seurObj, reduction='pca_CC', shuffle=T, group.by = 'Phase', split.by='sample') + labs(title='After cycle correction')

p1
p2
```


```{r}
SaveH5Seurat(seurObj, "GEX_cellcycle.h5seurat", overwrite = TRUE)
```

